---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: Analysis of Ebola data from WHO
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
bibliography: 
biblio-style: apsr
endnote: no
---




```{r}

WHO_SL <- WHO_bydistricts %>%
              filter(Country == "Sierra Leone") %>%
              droplevels %>% na.omit


ggplot(WHO_SL, aes(Date, incid)) +
    geom_point() +
    facet_wrap(~CL_DistrictRes, ncol = 3) +
    theme_minimal() +
    ggtitle("Incidence in the districts of Sierra Leone")

```


We will proceed as we did with HealthMap data - split the data set into training and
validation sets, set up the parameters and press play.

```{r}
SL_wide <-  WHO_SL %>%
            ungroup %>%
            select(-Country) %>%
            tidyr::spread(CL_DistrictRes, incid, fill = 0)
               

```

We are now ready to estimate R and move forward with the projection
exactly as we did for HealthMap data. 

```{r}

start     <- 2:(length(SL_wide$Date) - time_window)
end       <- start + time_window
end.dates <- SL_wide[end, "Date"]
r.estim   <- SL_wide  %>%
                   select(-Date) %>%
    plyr::alply(2, .dims = TRUE, function(incid) {
                                   I   <- pull(incid, 1) 
                                   res <- EstimateR(I, T.Start = start , T.End = end,
                                                    method = "NonParametricSI",
                                                    SI.Distr = SI_Distr,
                                                    plot = FALSE ,
                                                    CV.Posterior = 1 ,
                                                    Mean.Prior = 1 ,
                                                    Std.Prior = 0.5)
                                   res$R %<>% cbind(Date = end.dates)
                                   return(res$R)})
r.estim %>%
    bind_rows(.id = "District") %>%
    ggplot(aes(Date, `Mean(R)`)) +
    geom_point() +
    facet_wrap(~District) +
    ggtitle("Mean R for each district in Sierra Leone")

```

Next, calculate the population flow matrix from gravity model and
hence the relative risk for each district.



```{r}
pow_N_from <- 1
pow_N_to   <- 1
pow_dist   <- 2
K          <- 1

adm2_centroids <- "data/Geography/GravityModel/raw/adm2.txt" %>%
                   read.csv(stringsAsFactors = FALSE,
                            sep = "\t",
                            header = TRUE) %>%
                   filter(ADM0 == "Sierra Leone")  %>%
                   filter(ADM2 != "Western Rural") 

adm2_centroids$ADM2 %<>% gsub("Western Urban", "Western", .)
adm2_centroids$ADM2  %<>%
    gsub(' ', '', .) %<>%
    toupper %<>%
    factor

levels(adm2_centroids$ADM2) <- levels(WHO_SL$CL_DistrictRes)
adm2_centroids %<>% arrange(ADM2)
 
flow.matrix  <- flow_matrix(longitude   = adm2_centroids[, "Centroid_Lon"],
                            latitude    = adm2_centroids[, "Centroid_Lat"],
                            population  = adm2_centroids[, "Pop"],
                            place.names = adm2_centroids[, "ADM2"],
                            model = "gravity", K = K,
                            pow_N_from = pow_N_from,
                            pow_N_to   = pow_N_to,
                            pow_dist   = pow_dist)


## Relative risk
relative.risk <- flow.matrix / rowSums(flow.matrix, na.rm=TRUE)
p.stay     <- 0.99 
p.movement <- probability_movement(relative.risk, p.stay)

```

At this point, all the pieces are in place.
SL_training contains the incidence count we will use for projecting.
Estimate of R is in the data frame r.j.t with a row for each
simulation we want to run. p.movement conatins the probabilities.
We will project over a period of 6 weeks at three
different time points - at the start of the epidemic, during the
middle and towards the end.

## Projections

```{r}
n.sim  <- 1000L    # Number of simulations to run
t.proj <- 250L
r.j.t  <- r.estim %>%
           lapply(function(R){
                    cutoff <- which(R$Date %in% SL_wide[t.proj, "Date"])
                    shape  <- R[cutoff, "Mean(R)"]^2 / R[cutoff, "Std(R)"]^2
                    scale  <- R[cutoff, "Std(R)"]^2 / R[cutoff, "Mean(R)"]
                    return(rgamma(n.sim, shape = shape,
                                  scale = scale))}) %>%
            data.frame(check.names = F)

n.dates.sim   <- 49L      # The time period over which projection will be made.
SL_training   <- SL_wide[1:t.proj, ]
validation    <- nrow(SL_wide)
SL_validation <- SL_wide[(1 + t.proj):validation, ]

incid     <- as.matrix(select(SL_training, -Date))
dates.all <- SL_training[1:t.proj, ] %>%
             pull(Date) %>%
             c(seq(max(.) + 1, length.out = n.dates.sim, by = 1))
t.max     <- t.proj + n.dates.sim - 1


daily.projections <- plyr::alply(r.j.t, 1, function(r.t){
                                    r.t   <- as.matrix(r.t)
                                    out   <- project(incid, r.t, SI_Distr,
                                                     p.movement, n.dates.sim) 
                                    incidence.proj <- rbind(incid, out)
                                    incidence.proj %<>% data.frame(check.names = F)
                                    incidence.proj %<>% cbind(Date = dates.all)
                                    return(incidence.proj[(nrow(SL_training) + 1):t.max, ])})


weekly.projections <- lapply(daily.projections, daily.to.weekly) %>% bind_rows(.)

```
At the same time, we will also obtain the weekly incidence count from
available data.

```{r, warning=FALSE}
weekly.available <- c(training    = list(SL_training),
                       validation = list(SL_validation)) %>%
                       lapply(daily.to.weekly) %>%
                       bind_rows(.id = "Category")

end <- max(dates.all) + 42
weekly.available %<>% filter(Date < end)
plots.list <- SL_training %>%
                 select(-Date) %>%
                 colnames %>%
                 lapply(function(location){
                         available  <- weekly.available %>%
                                       `[`(, c("Date", "Category", location))
                         projection <- weekly.projections %>%
                                       `[`(, c("Date", location))
                         plot.weekly(available, projection)})

cowplot::plot_grid(plots.list[[1]],
                        plots.list[[2]],
                        plots.list[[3]],
                        plots.list[[4]],
                        plots.list[[5]],
                        plots.list[[6]],
                        plots.list[[7]],
                        plots.list[[8]],
                        plots.list[[9]],
                        plots.list[[10]],
                        plots.list[[11]],
                        plots.list[[12]],
                        plots.list[[13]])                   
 

```

## Visualising the projections on a map

We will only plot the incidence data from 1 week before we started
projection to the end of the projection period.


```{r}

sierra         <- rgdal::readOGR(dsn = "data/Geography/SLE_adm_shp",
                                 layer = "SLE_adm2")
sierra@data$id <- sierra@data$NAME_2
sierra.points  <- broom::tidy(sierra, region="id")
sierra.points$id %<>% factor
sierra.df      <- left_join(sierra.points, sierra@data, by="id")

sierra.df$NAME_2 %<>%
    toupper %<>%
    gsub(' ', '', .) %<>%
    factor


start           <- SL_wide$Date[t.proj] - 7
end             <- max(dates.all) 
available_small <- filter(weekly.available, Date > start & Date < end) %>% select(-Category)
projection_50   <- weekly.projections %>% group_by(Date) %>% summarise_all(quantile, probs = 0.5)
weekly_incid    <- list(training = available_small, projection = projection_50) %>% bind_rows(.id = "Category")
library(lubridate)
weekly_incid$week <- interval(start, weekly_incid$Date) %/% lubridate::weeks(1)
weekly_incid    %<>% tidyr::gather(key = NAME_2, value = incid, factor_key = TRUE, -c(Date, week, Category))
## to get rid of 0s as we will take log later.
weekly_incid$incid %<>% `+`(1)
sierra.df$NAME_2 <- plyr::revalue(sierra.df$NAME_2,
                                  c("WESTERNURBAN" = "WESTERN",
                                    "WESTERNRURAL" = "WESTERN"))

incid_map  <- left_join(sierra.df, weekly_incid)

incid_map$Category %<>% factor(levels = c("training", "projection"))
incid_map %<>% arrange(Category)


 ggplot(incid_map) + 
    aes(long, lat, group = group, fill = incid) +
    geom_polygon() +
    geom_path(color = "white" ) +
    coord_equal() +
    facet_grid(Category~week) +
    scale_fill_gradient2(low = "grey50",
                         high = "firebrick",
                         trans = "log10")
                         


```
