---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: 
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---



First read in and clean up the data from Health Map. Since these data
are at national level, we will also aggregate the WHO data at country
level. HealthMap collects and curates data on the cumulative number
of cases (C) and deaths (D) stratified by status (suspected, SC/SD, or
confirmed CC/CD). For data from the WHO, the case status could be
confirmed, probable or suspected. The final clinical outcome (alive/dead)
for each case is also recorded. In comparing the two data sets, we sum
across all four categories in Healthmap and derive incidence data from
cumulative data.

```{r hmread, echo = FALSE}
species   <- "Humans"
disease   <- "Ebola"
case.type <- "SCC"
healthmap <- "data/CaseCounts/raw/HealthMap_Ebola_GNE_WHO.csv" %>%
               read.csv(stringsAsFactors = FALSE) %>%
               filter(species == species,
                             disease == disease,
                             case.type == case.type,
                             Country %in% unique(WHO_raw$Country) )

healthmap_bycountry <- split(healthmap, healthmap$Country) 

healthmap_weekly    <- healthmap_bycountry %>%
                       lapply(function(case.count){
                               location <- case.count$Country[1]
                               case.count %<>%
                                  incidence.from.DS1(species, disease,
                                                   case.type,
                                                   location,
                                                   merge_rule = "median") %<>%
                                 daily.to.weekly }) %>%
                        bind_rows(.id = "Country")

healthmap_weekly$Date %<>% lubridate::ymd(.)

```

ProMed data is similarly stratfied and we will sum across the
catgories of suspected and confirmed cases for ProMed.

```{r pmread, echo=FALSE}

promed <- read.csv("data/CaseCounts/raw/ProMED_Ebola_2014-2016.csv",
                   stringsAsFactors = F)
promed$Issue.Date %<>% lubridate::dmy_hm(.)
promed$Issue.Date %<>% format(format = "%m/%d/%y %H:%M")

promed %<>% rename("SC" = "Cumulative.SC",
                   "SD" = "Cumulative.SD",
                   "CC" = "Cumulative.CC",
                   "CD" = "Cumulative.CD",
                   "HealthMap.Alert.ID" = "HM.Alert.ID")

promed    %<>% filter(species == species,
                      disease == disease,
                      Country %in% unique(WHO_raw$Country) )

promed_bycountry <- split(promed, promed$Country) 

promed_weekly    <- promed_bycountry %>%
                       lapply(function(case.count){
                               location <- case.count$Country[1]
                               case.count %<>%
                                  incidence.from.DS1(species, disease,
                                                   case.type,
                                                   location,
                                                   merge_rule = "median") %<>%
                                 daily.to.weekly }) %>%
                        bind_rows(.id = "Country")

promed_weekly$Date %<>% lubridate::ymd(.)

```


```{r who_hm_pm_weekly_incid, eval = TRUE}
WHO_weekly <- split(WHO_bycountry, WHO_bycountry$Country) %>%
              lapply(function(df){
                      df      %>%
                      ungroup %>%
                      select(Date, incid) %>%
                      daily.to.weekly}) %>%
                bind_rows(.id = "Country")    
WHO_weekly$Date       %<>% lubridate::ymd(.)


color_scale <- c("HealthMap" = "black", "WHO" = "green", "ProMed" = "red")

p <- list( WHO = WHO_weekly,
     HealthMap = healthmap_weekly[, c("Country", "Date", "incid")],
     ProMed    = promed_weekly[, c("Country", "Date", "incid")]) %>%
    bind_rows(.id = "Source") %>%
    ggplot(aes(Date, incid, color = Source)) +
    geom_point(size = 0.8) +
    facet_grid(Country ~.) +
    scale_colour_manual(values = color_scale)


p <- p + theme_bw() +
    theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"),
          strip.background = element_blank())
p <- p + theme(axis.text.x = element_text(angle = 80, hjust = 1))
p <- p + theme(legend.title = element_blank())
p <- p + xlab("")
p <- p + scale_x_date(date_labels =  "%b %Y")
p



```
## Estimating Reproduction number from WHO and Healthmap Data

```{r who_vs_hm_vs_pm}

promed_R <- promed_bycountry %>%
               lapply(function(case.count){
                        location <- case.count$Country[1]
                        case.count %<>%
                        incidence.from.DS1(species, disease,
                                           case.type,
                                           location,
                                           merge_rule = "median")
                        start     <- 1:(length(case.count$Date) - time_window)
                        end       <- start + time_window
                        end.dates <- case.count$Date[end] 
                        res       <- EstimateR(case.count$incid,
                                               T.Start  = start ,
                                               T.End    = end,
                                               method   = "NonParametricSI",
                                               SI.Distr = SI_Distr,
                                               plot     = FALSE ,
                                               CV.Posterior = 1 ,
                                               Mean.Prior   = 1 ,
                                               Std.Prior    = 0.5)
                        res$R %<>% cbind(Date = end.dates)
                        return(res$R)}) %>%
              bind_rows(.id = "Country")

healthmap_R <- healthmap_bycountry %>%
               lapply(function(case.count){
                        location <- case.count$Country[1]
                        case.count %<>%
                        incidence.from.DS1(species, disease,
                                           case.type,
                                           location,
                                           merge_rule = "median")
                        start     <- 1:(length(case.count$Date) - time_window)
                        end       <- start + time_window
                        end.dates <- case.count$Date[end] 
                        res       <- EstimateR(case.count$incid,
                                               T.Start  = start ,
                                               T.End    = end,
                                               method   = "NonParametricSI",
                                               SI.Distr = SI_Distr,
                                               plot     = FALSE ,
                                               CV.Posterior = 1 ,
                                               Mean.Prior   = 1 ,
                                               Std.Prior    = 0.5)
                        res$R %<>% cbind(Date = end.dates)
                        return(res$R)}) %>%
              bind_rows(.id = "Country")

WHO_R <- WHO_bycountry    %>%
         split(.$Country) %>%
         lapply(function(case.count){
                  location  <- case.count$Country[1]
                  start     <- 1:(length(case.count$Date) - time_window)
                  end       <- start + time_window
                  end.dates <- case.count$Date[end] 
                  res       <- EstimateR(case.count$incid,
                                         T.Start  = start ,
                                         T.End    = end,
                                         method   = "NonParametricSI",
                                         SI.Distr = SI_Distr,
                                         plot     = FALSE ,
                                         CV.Posterior = 1 ,
                                         Mean.Prior   = 1 ,
                                         Std.Prior    = 0.5)
                  res$R %<>% cbind(Date = end.dates)
                  return(res$R)}) %>%
           bind_rows(.id = "Country")


p <- list( WHO = WHO_R, HealthMap = healthmap_R, ProMed = promed_R) %>%
    bind_rows(.id = "Source") %>%
    ggplot(aes(Date, `Mean(R)`, color = Source)) +
    geom_line() +
    facet_grid(Country ~.)

p <- p + theme_bw() +
    theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(colour = "black"),
          strip.background = element_blank())
p <- p + theme(axis.text.x = element_text(angle = 80, hjust = 1))
p <- p + theme(legend.title = element_blank())
p <- p + xlab("")
p <- p + scale_x_date(date_labels =  "%b %Y")
p <- p + scale_colour_manual(values = color_scale)
p <- p + ggtitle("Mean R estimated from incidence data") +
         theme(plot.title = element_text(hjust = 0.5))

p
```

The graphs show that the estimates from the three different sources
closely track each other. Here's another way to visualise this.

```{r hm_who}

HM_vs_WHO <- inner_join(WHO_R, healthmap_R,
                        by = c("Date", "Country")) %>%
              select(Country, Date, `Mean(R).x`,  `Mean(R).y`) %>%
              rename("WHO" = `Mean(R).x`,
                     "HealthMap" = `Mean(R).y`)

ggplot(HM_vs_WHO, aes(HealthMap, WHO)) +
    geom_point() +
    facet_wrap(~Country, nrow = 3) +
    xlab("R estimated from HealthMap data") +
    ylab("R estimated from WHO data") +
    geom_abline(slope = 1)
```

```{r pm_who}

promed_vs_WHO <- inner_join(WHO_R, promed_R,
                        by = c("Date", "Country")) %>%
              select(Country, Date, `Mean(R).x`,  `Mean(R).y`) %>%
              rename("WHO" = `Mean(R).x`,
                     "Promed" = `Mean(R).y`)

ggplot(promed_vs_WHO, aes(Promed, WHO)) +
    geom_point() +
    facet_wrap(~Country, nrow = 3) +
    xlab("R estimated from ProMed data") +
    ylab("R estimated from WHO data") +
    geom_abline(slope = 1)


```

```{r pm_hm}

promed_vs_hm <- inner_join(healthmap_R, promed_R,
                        by = c("Date", "Country")) %>%
              select(Country, Date, `Mean(R).x`,  `Mean(R).y`) %>%
              rename("HealthMap" = `Mean(R).x`,
                     "ProMed" = `Mean(R).y`)

ggplot(promed_vs_hm, aes(HealthMap, ProMed)) +
    geom_point() +
    facet_wrap(~Country, nrow = 3) +
    xlab("R estimated from HealthMap data") +
    ylab("R estimated from ProMed data") +
    geom_abline(slope = 1)

```
As the above graphs indicates, estimates of the reproduction number
from ProMed, HealthMap and WHO data are highly correlated (correlation
coefficient = 0.76, 95% CI = (0.74, 0.78)).

```{r}
hm_who_pm <- inner_join(WHO_R, healthmap_R,
                        by = c("Date", "Country")) %>%
              select(Country, Date, `Mean(R).x`,  `Mean(R).y`) %>%
              rename("WHO" = `Mean(R).x`,
                     "HealthMap" = `Mean(R).y`) %>%
              inner_join(promed_R, by = c("Date", "Country")) %>%
              rename("ProMed" = `Mean(R)`)  

d <- data.frame(ProMed    = hm_who_pm$ProMed,
                WHO       = hm_who_pm$WHO,
                HealthMap = hm_who_pm$HealthMap) %>% na.omit

cor(d)
#cor.test(HM_vs_WHO$WHO, HM_vs_WHO$HealthMap)
```
