\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{natbib}
\usepackage[margin=3cm]{geometry}
\usepackage{todonotes}
\usepackage{mathtools}
\usepackage{subfiles}
\usepackage{subcaption}
\usepackage{booktabs}
\def\sb#1{\textcolor{blue}{#1}}

\begin{document}

\begin{center}
 \textbf{\large Responses to Review Comments}
\end{center}

\sb{Responses to all the comments have been incorporated in the report and the
edited/additional text has been highlighted in bold.}

\begin{itemize}
\item[] Page 10: Meaning the data is aggregated at a national level? What does this mean during a crisis? What will we be able to know/not know?
 

\sb{Comment : Yes, the data are aggregated at national level. This means that analysis cannot be carried out at a finer spatial scale. For instance, we cannot do forecasting at a sub-national level or understand how an epidemic is spreading within a country.}
 

\item[] Page 10: But this can only be done after the fact/ after a crisis?

\sb{Comment : This analysis can be carried out as and when the data from sub-national level starts becoming available. We are still experimenting with this idea but in principle, it would be possible to use historical data to carry out this analysis.}
 

\item[] Page 11: Can you extrapolate a bit more here? 

\sb{Comment: The strength of the correlation is indicated by the size and colour of the circles. Deeper shades of blue indicate high correlation. The figure shows that the data collected by ProMed and HealthMap exhibits a high degree of correlation with the data collected by WHO indicating that ProMed/HealthMap data are reliable.}
 

\item[] Page 16: Can you be more direct here? What did the model accomplish?
Did it predict things well? Could it be done in real time, given the
data one would be able to get in real time? Are there shortcomings that could/should be filled by other data sources? If so, have you considered what those maybe?
 
\sb{Comment: We have included detailed comments on assessing the fit of
the model in Section 4.6.2. The section also includes our comments on
the other questions about the real-time nature of the analysis and
shortcomings of the current model.}
\end{itemize}
\end{document}