---
output: 
pdf_document:
citation_package: natbib
keep_tex: true
fig_caption: true
latex_engine: pdflatex
title: Pull out centroids for regions in Central Africa
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---

```{r setup, eval = TRUE}
library(dplyr)
library(stringr)
library(ggplot2)
```

For some countries we want to deal at ADM1 level while for others we
want to work at ADM2 level. 

If we choose to work at ADM2 level in DRC then we should work at ADM1 
in all other countries (if we try to match populations).
If, on the other hand, we work at ADM1 level in DRC then the closest
population match is achieved by working at the following levels.

|ADM0                             |level |
|:--------------------------------|:-----|
|Angola                           |ADM1  |
|Burundi                          |ADM0  |
|Central African Republic         |ADM0  |
|Democratic Republic of the Congo |ADM1  |
|Republic of Congo                |ADM0  |
|Rwanda                           |ADM0  |
|South Sudan                      |ADM0  |
|Tanzania                         |ADM1  |
|Uganda                           |ADM1  |
|Zambia                           |ADM0  |


For ease of analysis, pull out the
meta-data and put them together.


## Read in ADM2 data

```{r pop_match, eval = TRUE}
#adm2_neighbors <- c("Democratic Republic of the Congo")

adm1_neighbors <- c("Angola", "Burundi", 
                   "Uganda", "Republic of Congo",
                   "Central African Republic",
                   "South Sudan",
                   "Rwanda",
                   "Zambia")
                   

outfile <- here::here("data/Geography/GravityModel/processed",
                      "drc_neighbors_pops.csv")


```

## Working to match spatial area

```{r area_match, eval = TRUE}
adm1_neighbors <- c("Uganda",
                    "Angola",
                    "Central African Republic",
                    "Republic of Congo",
                    "South Sudan",
                    "Zambia")
##                    "Tanzania")

adm0_neighbors <- c("Burundi", 
                    "Rwanda")

outfile <- here::here("data/Geography/GravityModel/processed",
                      "drc_neighbors_area.csv")

```

## Stack them together

```{r}
centroids_adm2 <- here::here("data/Geography/GravityModel/raw",
                             "adm2-fixed.txt") %>%
    readr::read_tsv(.) %>%
    filter(ADM0 %in% adm2_neighbors)


centroids_adm1 <- here::here("data/Geography/GravityModel/raw",
                             "adm1_centroids_fixed.tsv") %>%
    readr::read_tsv(.) %>%
    filter(ADM0 %in% adm1_neighbors)

centroids_adm0 <- here::here("data/Geography/GravityModel/raw",
                             "adm0.txt") %>%
    readr::read_tsv(.) %>%
    filter(ADM0 %in% adm0_neighbors) %>%
    select(ADM0,
           Centroid_Lon,
           Centroid_Lat,
           Pop)    

centroids_adm0$district <- centroids_adm0$ADM0

centroids <- select(centroids_adm2,
                    ADM0,
                    district = ADM2,
                    Centroid_Lon,
                    Centroid_Lat,
                    Pop)

centroids <- select(centroids_adm1,
                    ADM0,
                    district = ADM1,
                    Centroid_Lon,
                    Centroid_Lat,
                    Pop) #%>% rbind(centroids)

centroids <- rbind(centroids, centroids_adm0)

                    
```
## Affix DRC Centroids.
```{r}
drc_metadata <- here::here("data/Geography/GravityModel/processed",
                           "drc_metadata.csv") %>%
    readr::read_csv() %>%
    select(ADM0,
           district = adm2,
           Centroid_Lon,
           Centroid_Lat,
           Pop = pop)

centroids <- rbind(centroids, drc_metadata)
```
## Write out processed file

```{r}


readr::write_csv(centroids,
                 path = outfile)
```

