---
output: 
pdf_document:
citation_package: natbib
keep_tex: false
fig_caption: true
latex_engine: pdflatex
title: Choose appropriate ADM level for DRC neighbors
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---

```{r setup, eval = TRUE}
library(dplyr)
library(stringr)
library(ggplot2)
```
We want to choose for analysis the administrative subdivision of the 
neighbors of DRC that most closely matches the subdivisions of DRC in 
population or spatial area.
To do this, we determine the average population in the subdivisions of 
DRC and the average population at ADM0, ADM1 and ADM2 levels in its
neighbors and for each country, choose the subdivision that is closest.

## Read in data for DRC
```{r}
drc_metadata <- here::here("data/Geography/GravityModel/processed",
                           "drc_metadata.csv") %>%
    readr::read_csv()

drc_mean_area <- mean(drc_metadata$area)
drc_mean_pop <- mean(drc_metadata$pop) ## use the one we calculated.
```
## List countries we are interested in

```{r}
drc_neighbors <- c("Angola",
                   "Burundi",
                   "Central African Republic",
                   "Republic of Congo",
                   "Rwanda",                   
                   "South Sudan",
                   "Tanzania",
                   "Uganda",
                   "Zambia")

```

## Read in ADM1 data

```{r}
centroids_adm1 <- here::here("data/Geography/GravityModel/raw",
                             "adm1_centroids_fixed.tsv") %>%
    readr::read_tsv(.) %>%
    filter(ADM0 %in% drc_neighbors)

mean_pop_adm1 <- group_by(centroids_adm1, ADM0) %>%
    summarise(mean_adm1 = mean(Pop)) 
```

## Read in ADM2 data

```{r}
centroids_adm2 <- here::here("data/Geography/GravityModel/raw",
                             "adm2-fixed.txt") %>%
    readr::read_tsv(.) %>%
    filter(ADM0 %in% drc_neighbors)

mean_pop_adm2 <- group_by(centroids_adm2, ADM0) %>%
    summarise(mean_adm2 = mean(Pop)) 

pop_adm0 <- group_by(centroids_adm2, ADM0) %>%
    summarise(pop_adm0 = sum(Pop)) 
```

## Put them together

```{r}
mean_pops <- left_join(pop_adm0, mean_pop_adm1) %>%
    left_join(mean_pop_adm2)
```

## Choose wisely

For each country, choose the level that most closely matches the 
level at which you want to operate in DRC.

```{r}


## From vector "among" choose the index closest to
## to
choose_closest <- function(among, to) {
    delta <- abs(among - to)
    idx <- which(delta == min(delta))
    idx
}

    
    
nghbr_levels <- c()
for(nghbr in drc_neighbors) {
    among <- filter(mean_pops, ADM0 == nghbr) %>% `[`(, 2:4)
    idx <- choose_closest(among, drc_mean_pop)
    idx <- paste0("ADM", idx - 1)
    nghbr_levels <- c(nghbr_levels, idx)
}    

nghbr_levels_pops <- data.frame(ADM0 = drc_neighbors, level = nghbr_levels)
```

```{r}
knitr::kable(nghbr_levels_pops)
```

## Choose according to area

```{r}
mean_areas <- here::here("data/Geography",
                         "drc_neighbors_areas.csv") %>%
    readr::read_csv()



drc_neighbors <- mean_areas$ADM0
nghbr_levels <- c()
for(nghbr in drc_neighbors) {
    among <- filter(mean_areas, ADM0 == nghbr) %>% `[`(, 2:4)
    idx <- choose_closest(among, drc_mean_area)
    idx <- paste0("ADM", idx - 1)
    nghbr_levels <- c(nghbr_levels, idx)
}    

nghbr_levels_area <- data.frame(ADM0 = drc_neighbors, level = nghbr_levels)
```

```{r}
nghbr_levels_area <- arrange(nghbr_levels_area, level)
knitr::kable(nghbr_levels_area)
``
