---
output: 
pdf_document:
citation_package: natbib
keep_tex: true
fig_caption: true
latex_engine: pdflatex
title: Cumulative incidence into daily incidence
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  cases: 14_to_17_may.csv
  risk: !r c("flow_from_mbandaka_2.csv", "flow_from_Équateur_2.csv")
  pstay: 0.95
  sim: 10000
  simean: 16
  sisd: 4.4
---

```{r setup, eval = TRUE}
library(dplyr)
library(magrittr)
library(stringr)
library(ggplot2)
devtools::load_all()
```
This file outputs an incidence time series from cumulative case count.

##  Case Counts
Read in the case counts from the sitrep and calculate overall
infectivity at the current time.

```{r}
cum_cases <- here::here("data/CaseCounts/drc",
                        params$cases) %>%
    readr::read_csv()

cum_cases$date <- lubridate::dmy(cum_cases$date)
```
## Fixing data

According to the report on 17th May, the cumulative case count in
Bikoro has *reduced* from 36 to 29. Ignore this point.

```{r}
cum_cases <- cum_cases[ -nrow(cum_cases), ]
```

Assume a starting point.

```{r, eval = FALSE}
cum_cases <- rbind(data.frame(date = lubridate::dmy("01/05/20180"),
                              Bikoro = NA,
                              Iboko = NA,
                              Wangata = NA),
                   cum_cases)

```
## Interpolation

Fit a line to fill in missing data.

```{r}
linear_fit <- select_if(cum_cases, is.numeric) %>%
    purrr:: map_dfr(function(x) {
        interpolate_missing_data(cum_incidence = data.frame(date = cum_cases$date,
                                                            cases = x),
                                 method = "linear")},
        .id = "district")

linear_fit <- select(linear_fit,
                     district,
                     interpolated_date,
                     interpolated_cases) %>%
    tidyr::spread(district, interpolated_cases)
```


BIKORO and Iboko are in Equateur province while Wangata is in Mbandaka
province.

```{r}
linear_fit$eq_cases <- linear_fit$Bikoro + linear_fit$Iboko
linear_fit$mb_cases <- linear_fit$Wangata
```

And the incidence curve from the cumulative case count.
```{r}
eq_incid <- c(linear_fit$eq_cases[1], diff(linear_fit$eq_cases))
mb_incid <- c(linear_fit$mb_cases[1], diff(linear_fit$mb_cases))
```

## Write out incidence curve

```{r}
here::here("data/CaseCounts/drc",
           paste0("interpolated_", params$cases)) %>%
    readr::read_csv(x = linear_fit, path = .)
```
