---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Borrowing information across spatial scales 
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  n.sim : 1000
  ADM0: SIERRALEONE
  bin_from: 2015-W02
---

```{r global_options, include=FALSE}

knitr::opts_chunk$set(
  fig.width = 12, fig.height = 6,
  echo = FALSE,
  warning = FALSE, message = FALSE,
  fig.path = "figures/"
)
```

```{r setup, eval = TRUE}

library(dplyr)
library(stringr)
library(ggplot2)
library(purrr)
library(EpiEstim)
devtools::load_all()
```

## Unpack parameters

```{r pars_read, eval = TRUE}

n.sim <- params$n.sim
adm0 <- params$ADM0
bin_from <- params$bin_from
```

## Model

We will assign the national case count to districts according to a
multinomial distribution with the probabilities derived from gravity
model or simply population density. 

First we read in the data that is required by both methods.

## Ebola Parameters
```{r ebola_params}

mean_SI <- 14.2
CV_SI <- 9.6 / 14.2
SItrunc <- 40
SI_Distr <- sapply(0:SItrunc, function(e) DiscrSI(e, mean_SI, mean_SI * CV_SI))
SI_Distr <- SI_Distr / sum(SI_Distr)
```


## Incidence data from HealthMap
This is the aggregated data that will be used to assign cases to
districts.

```{r hm_data, eval = TRUE}

hm_wide <- here::here(
  "data/CaseCounts/processed",
  "HealthMap_Ebola_wide.csv"
) %>%
  readr::read_csv(.)
```  

We will work with weekly incidence counts. 

```{r hm_weekly, eval = TRUE}

hm_weekly <- daily.to.weekly(hm_wide)
```
Also add isoweek column so that we can compare across the three data
sets.

```{r}
hm_weekly$week_of_year <- paste0(lubridate::year(hm_weekly$Date),
                            "-W",
                            stringr::str_pad(lubridate::isoweek(hm_weekly$Date),
                                             2,
                                             pad = "0"))

hm_weekly$week_of_year <- factor(hm_weekly$week_of_year, ordered = TRUE)
```
## Weekly Training Incidence Data 

```{r, eval =TRUE}
who_wide <- here::here("data/CaseCounts/processed",
                       "who_sitrep_wide_sl_2016-05-11.csv") %>%
    readr::read_csv()
who_wide$week_of_year <- factor(who_wide$week_of_year, ordered = TRUE)
```


## Validation data

The performance of the model will be validated against the cleaned-up
data.


```{r who, eval = TRUE}

sl_weekly <- here::here(
  "data/CaseCounts/processed",
  "sl_weekly_tall.csv"
) %>%
  readr::read_csv(.) 

sl_weekly$week_of_year <- paste0(lubridate::year(sl_weekly$Date),
                            "-W",
                            stringr::str_pad(lubridate::isoweek(sl_weekly$Date),
                                      2,
                                      pad = "0"))
sl_weekly$week_of_year <- factor(sl_weekly$week_of_year, ordered = TRUE)
```

## Further cleaning
Maybe get rid of PUJEHUN as it has low incidence (difficult to fit any
model)

```{r, eval = FALSE}
who_wide <- select(who_wide, -PUJEHUN)  
sl_weekly <- filter(sl_weekly, district != "PUJEHUN")

```

## Method 1: Relative to the reproduction number from recent past

### 3A: Use data published by WHO

Suppose we are doing the disaggregation in the last week of 2014.
Read in the latest information published by WHO.
Start binning at date that ends week in variable ``bin_from''.
It might be better to use recent data rather than all data
available upto the point of projection.

```{r}
use_last <- 6 ## weeks data
trng_end_level <- which(levels(who_wide$week_of_year) == bin_from)
trng_start_level <- if (trng_end_level < (use_last + 1)) 1 else trng_end_level - use_last
trng_start <- levels(who_wide$week_of_year)[trng_start_level]
trng_end <- bin_from
training <- filter(
  who_wide,
  week_of_year >= trng_start,
  week_of_year <= trng_end
)
```

## Fitting log-linear model

Add 1 to avoid incidence::fit rejecting all 0s and throwing an error.
Dubious but lets go with it for now.

```{r}
ll_fit <- select_if(training, is.numeric) %>%
    map(function(I){
        message(I)
        I <- I + 1
        incid <- incidence::as.incidence(x = I,
                                         interval = 7,
                                         isoweek = TRUE)
        incidence::fit(incid)
    })

district_rates <- map(ll_fit, function(fit){
    epitrix::r2R0(fit$info$r, SI_Distr)})

prob <- matrix(unlist(district_rates), nrow = 1)
colnames(prob) <- colnames(training)[-1]

```

## Assign cases to districts

```{r write_probs, eval = FALSE}

prob_df <- data.frame(prob) %>% tidyr::gather(district, prob)
prob_df$rel_prob <- prob_df$prob / sum(prob_df$prob)
prob_df <- arrange(prob_df, district)
paste0(names(method)[which(method == TRUE)], "_probs.csv") %>%
  readr::write_csv(x = prob_df, path = .)
```

Extract the data we want to use for disaggreation.

```{r}
bin_for <- 6 ## weeks
num_weeks <- length(levels(hm_weekly$week_of_year))
start_at <- which(levels(hm_weekly$week_of_year) == bin_from)
end_at <- min(num_weeks, start_at + bin_for)


bin_end   <- levels(hm_weekly$week_of_year)[end_at]

bin_this <- filter(
  hm_weekly,
  week_of_year >= bin_from,
  week_of_year <= bin_end
) %>%
    droplevels() %>%
    select(week_of_year, `Sierra Leone`)
```

						 
```{r assign, eval = TRUE}
district_weekly <- split(bin_this, bin_this$week_of_year) %>%
  lapply(function(df) {
    mat <- disaggregate(
      total = df$`Sierra Leone`,
      pmatrix = prob
    )
    colnames(mat) <- colnames(prob)
    probs <- c(0.025, 0.5, 0.975)
    distr <- disaggregate_distr(mat, probs = probs)
    distr
  }) %>%
  bind_rows(.id = "week_of_year")

district_weekly$bin_from <- bin_from

```

Write out the output. 

```{r}
here::here("output", paste0(
 "projections_",                        
  bin_from,
  ".csv"
)) %>%
    readr::write_csv(x = district_weekly, path = .)

```
Also write out the training data.

## Training Data
```{r}
who_tall <- filter(who_wide, week_of_year >= trng_start &
                             week_of_year <= bin_end) %>%
    tidyr::gather(district, incid, -week_of_year)

```


```{r trng, eval = TRUE}


here::here("output", paste0(
  "training_",                         
  bin_from,
  ".csv"
)) %>%
    readr::write_csv(x = who_tall, path = .)



```

## Visualizing the results

## Validation data


```{r}
compare <- left_join(bin_this, sl_weekly)
compare$week_of_year <- factor(compare$week_of_year)
district_weekly$week_of_year <- factor(district_weekly$week_of_year)
district_weekly$district <- factor(district_weekly$district)
## Plot training data
p <- ggplot(who_tall, aes(week_of_year, incid)) + geom_point(col = "red")
p <- p + facet_wrap(~district, scale = "free_y")
## Plot validation data
p <- p + geom_point(data = compare, aes(x = week_of_year, y = incid))
## Plot projections
p <- p + geom_line(data = district_weekly,
                   aes(x = week_of_year, y = `0.5`, group = 1),
                   col = "blue")
p <- p + geom_ribbon(data = district_weekly,
                     aes(x = week_of_year,
                         ymin = `0.025`,
                         ymax = `0.975`,
                         group = 1),
                     alpha = 0.5,
                     inherit.aes = FALSE)
p <- p + theme_classic()
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p
here::here("output", paste0(bin_from, ".png")) %>% ggsave(p)
```
