---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: Comparison of data sources for healthcare capacity
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---

# World Bank
## [Hospital Beds per 1000 people](https://data.worldbank.org/indicator/SH.MED.BEDS.ZS)

Hospital beds include inpatient beds available in public, private,
general, and specialized hospitals and rehabilitation centers. In most
cases beds for both acute and chronic care are included. Annual.

```{r}
num_beds <- read.csv("data/HealthCareCapacity/API_SH.MED.BEDS.ZS_DS2_en_csv_v2.csv",
                      skip = 4)
colnames(num_beds)[5:61] <- 1960:2016
apply(num_beds[, 5:62], 2, function(col) sum(!is.na(col)))

```
Quite patchy. Out of the 264 countries for which data are recorded,
the most complete record is for the year 1990 when data for 187
countries are available. For 2012, only 70 countries have complete
records. In particular for West Africa, we have non-null records for
all countries for 1960 and for 5 of the 6 for 1975.

```{r}
w.africa <- c("Sierra Leone", "Guinea", "Liberia", "Mali", "Nigeria", "Ghana")
num_beds %>%
    filter(Country.Name %in% w.africa) %>%
    `[`(, 5:62) %>%
    apply(2, function(col) sum(!is.na(col)))

```
## [Number of physicians per 1000 people](https://data.worldbank.org/indicator/SH.MED.PHYS.ZS)

Physicians include generalist and specialist medical
practitioners. The WHO estimates that at least 2.5 medical staff
(physicians, nurses and midwives) per 1,000 people are needed to
provide adequate coverage with primary care interventions (WHO, World
Health Report 2006).

```{r}
physicians <- "data/HealthCareCapacity/NumberOfPhysicians/API_SH.MED.PHYS.ZS_DS2_en_csv_v2.csv" %>% read.csv(skip = 4)

colnames(physicians)[5:61] <- 1960:2016
apply(physicians[, 5:62], 2, function(col) sum(!is.na(col))) 

physicians %>%
    filter(Country.Name %in% w.africa) %>%
    `[`(, 5:62) %>%
    apply(2, function(col) sum(!is.na(col)))
```
Pretty much the same comments as before except that we have data for 5
    of the 6 countries for 2008.
	
# [IHR Core Capacities](http://apps.who.int/gho/data/node.main.IHR00ALLN?lang=en)
	
## [Hospital Beds Density](http://apps.who.int/gho/data/view.main.HS07v)
```{r}
ihr_beds <- "data/HealthCareCapacity/NumberOfBeds/WHS6_102.csv" %>%
             read.csv(skip = 1)
colnames(ihr_beds)[2:12] %<>% gsub("X.", "", .)
ihr_beds %>% apply(2, function(col) sum(!is.na(col)))
ihr_beds %>% filter(Country %in% w.africa)

```
Only Ghana and Guinea have data for 2011.

##[Physician Desnity](http://apps.who.int/gho/data/node.main.A1444?lang=en)

Grouped by the category of staff i.e., physicians, midwives etc. Also
quite complete for West Africa. For instance, the number of physicians
per 1000 people is available for all West African countries for all
years except for Nigeria in 2004 and 2011.

```{r}
ihr_physicians <- "data/HealthCareCapacity/NumberOfPhysicians/HRH_26,HRH_33,HRH_28,HRH_25,HRH_27,HRH_31,HRH_29,HRH_30,HRH_32.csv" %>% read.csv
ihr_physicians %>%
    filter(Country %in% w.africa) %>%
    apply(2, function(col) sum(!is.na(col)))

```

# Institute of Health Metrics and Evaluation
http://www.healthdata.org/results/country-profiles/haq
http://www.thelancet.com/journals/lancet/article/PIIS0140-6736(17)30818-8/abstract

This looks very good.

# Health Data Collaborative
https://www.healthdatacollaborative.org
