---
output: 
pdf_document:
citation_package: natbib
keep_tex: true
fig_caption: true
latex_engine: pdflatex
title: Relative Risk of Importation 
author:
- name: Sangeeta Bhatia
affiliation: Imperial College London
abstract: 
keywords: 
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  sources: !r c("Guinea", "Liberia")
  cases: data/CaseCounts/processed/HealthMap_Ebola_wide.csv
  risk: !r list(Guinea = "output/flow_from_mbandaka_2.01_72.96_1.12.csv", Liberia = "output/flow_from_équateur_2.01_72.96_1.12.csv")
  simean: 15.3
  sisd: 9.1
  R: 1.03
  onday: 300
  outfile: "output/weighted_flow.csv"
---
```{r setup, eval = TRUE, echo = FALSE}
library(dplyr)
library(purrr)
devtools::load_all()
```

This analysis uses a daily incidence time series to estimate the
overall infectivity at a given time. The relative risk profiles
estimated in a previous step of the analysis pipeline are then
weighted by infectivity.

## Ebola params
```{r}
incid <- here::here(params$cases) %>%
    readr::read_csv(n_max = params$onday) %>%
    select(date, params$sources)
```
```
Serial interval estimation over however many days worth of data we
have at this point.

```{r}
si <- EpiEstim::DiscrSI(k = 0:params$onday,
                        mu = params$simean,
                        sigma  = params$sisd) %>% round(3)

weights <- select(incid, params$sources) %>%
    map_dbl(~infectivity_at_source(.x, si, params$R) %>%
        `[`(params$onday))

```

# Normalise weights

```{r}
day <- params$onday
normalised <- weights / sum(weights)
```

Read in the relative risk profiles

```{r relrisk, eval = TRUE}

relrisk_df <- purrr::map(params$risk,
                         ~ readr::read_csv(here::here(.x)))

```

## Risk profile weighted by infectivity

```{r wtdrelrisk, eval = TRUE}
## signf <- 7
## normalised <- signif(normalised, 7)


## wtd_rel_risk <- (normalised[1]  * rel_risk[[1]]$relative_flow) +
##     (normalised[2] * rel_risk[[2]]$relative_flow)

rel_risk <- map_dfc(relrisk_df, ~ .x$relative_flow)
wtd_rel_risk <- map2_dfc(rel_risk,
                         normalised,
                         ~ .x * .y) %>%
    rowSums(na.rm = TRUE)

wtd_risk_profile <- data.frame(flow_to = relrisk_df[[1]]$flow_to,
                               wtd_rel_risk = wtd_rel_risk)
```

```{r wtdrelriskout, eval = TRUE}
readr::write_csv(x = wtd_risk_profile,
                 path = here::here(params$outfile))
```
