---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: Estimation of Model Parameters
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  pow_dist : 2
  t.proj : 300
  n.sim : 2
  n.dates.sim : 49
  p.stay: 0.9
  ADM0: GUINEA
---


```{r}
library(magrittr)
library(ggthemes)
library(ggplot2)
library(dplyr)
library(rstan)
library(EpiEstim)
devtools::load_all()

```

# Simple model in Stan

We start with defining a very simple model with only one 
location. 
The model is given by 
\[
 I(t) \sim Poisson(R_t \sum_{s = 0}^{t}{I_{t} ws_{t - s}}).
\]

The model is tested with simulated data.

# Parameter estimation for simulated data

Load in the data and prepare to feed into Stan.

```{r ebola_params, eval = TRUE}

mean_SI     <- 14.2
CV_SI       <- 9.6 / 14.2
SItrunc     <- 40
SI_Distr    <- sapply(0:SItrunc,
                      function(e) DiscrSI(e,
                                          mean_SI,
                                          mean_SI * CV_SI))
SI_Distr    <- SI_Distr / sum(SI_Distr)


```

## Example 1

```{r}

I0 <- 110
incid <- as.matrix(I0, 1, 1)
R <- c(rep(1.5, times = 8), rep(1.1, times = 7), rep(0.3, times = 7))
R <- as.matrix(R, ncol = 1)
pij <- 1 # probability of staying at i
n.days <- 21
SI <- SI_Distr[1:(n.days + 1)] 
I1 <- project2(incid, R, SI, pij, n.days)
I1 <- c(I0, I1)

```


And now fitting.

```{r}
sim_data <- list(T = length(I1),
                 I = I1,
                 SI = SI_Distr[1:length(I1)],
                 num_windows = 3,
                 windows_end = c(8, 15, length(I1)))

fit1 <- stan(
  file = "R/gravitymodel.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 1,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500          # show progress every 'refresh' iterations
)

print(fit1, pars=c("R"), probs=c(.1,.5,.9))

```
Looks like it is working as expected!

## Example 2

```{r}

I0 <- c(110, 150, 200, 300)
incid <- as.matrix(I0, length(I0), 1)
R <- c(rep(1.5, times = length(I0)),
           rep(1.5, times = 7),
           rep(1.1, times = 7),
           rep(0.3, times = 7))
R <- as.matrix(R, ncol = 1)
pij <- 1 # probability of staying at i
n.days <- 21
SI <- SI_Distr[1:(n.days + 1)] 
I1 <- project2(incid, R, SI, pij, n.days)
I1 <- c(I0, I1)

```


And now fitting.

```{r}
sim_data <- list(T = length(I1),
                 I = I1,
                 SI = SI_Distr[1:length(I1)],
                 num_windows = 3,
                 windows_end = c(12, 19, length(I1)))

fit1 <- stan(
  file = "R/gravitymodel.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 1,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500          # show progress every 'refresh' iterations
)

print(fit1, pars=c("R"), probs=c(.1,.5,.9))

```
Looks like it is working as expected!




# Parameter estimation for real data

```{r data_read, eval = TRUE}

infile     <- "WHO_bycountry.csv"
infile     <- here::here("data/CaseCounts/processed/", infile)
incid_wide <- readr::read_csv(infile)

guinea_incid <- filter(incid_wide, Country == "Guinea") %>%
                pull(incid)

lbr_incid <- filter(incid_wide, Country == "Liberia") %>%
                pull(incid)

sl_incid <- filter(incid_wide, Country == "Sierra Leone") %>%
                pull(incid)
 

```

## Using incidence from Guinea

Use the entire dataset.

```{r}

time_window <- 49
windows_end <- seq(time_window, length(guinea_incid), by = time_window)
num_windows <- length(windows_end)
T <- windows_end[num_windows]
I <- guinea_incid[seq(T)]

SI_len <- length(SI_Distr)
SI <- if(T < SI_len) SI_Distr[1:T] else c(SI_Distr, rep(0, T - SI_len))
  
sim_data <- list(T = T,
                 I = I,
                 SI = SI,
                 num_windows = num_windows,
                 windows_end = windows_end)




```

And get the ball rolling

```{r}

fit1 <- stan(
  file = "R/gravitymodel.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 5,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500          # show progress every 'refresh' iterations
)

print(fit1, pars=c("R"), probs=c(.1,.5,.9))
traceplot(fit1, pars = c("R"), inc_warmup = TRUE, nrow = 5)

from_stan <- summary(fit1)
from_stan <- from_stan$summary
from_stan <- from_stan[-13, c("mean", "sd", "2.5%", "25%",  "50%", "75%", "97.5%")]


```

## Estimate R using EpiEstim

```{r restim, eval = TRUE}
windows_start <- seq(1, by = time_window, length.out = length(windows_end))
r.estim   <- EpiEstim::EstimateR(I,
                                 T.Start = windows_start ,
                                 T.End = windows_end,
                                 method = "NonParametricSI",
                                 SI.Distr = SI_Distr,
                                 plot = FALSE ,
                                 CV.Posterior = 1 ,
                                 Mean.Prior = 1 ,
                                 Std.Prior = 0.5)

from_epiestim <- r.estim$R[, c("Mean(R)", "Std(R)",
                               "Quantile.0.025(R)",
                               "Quantile.0.25(R)",
                               "Median(R)",
                               "Quantile.0.75(R)",
                               "Quantile.0.975(R)")]

rownames(from_epiestim) <- rownames(from_stan)
colnames(from_epiestim) <- colnames(from_stan)


from_stan     <- tibble::rownames_to_column(data.frame(from_stan))
from_epiestim <- tibble::rownames_to_column(data.frame(from_epiestim))

```
And now compare them visually.

```{r}
compare <- list(stan = from_stan, epiestim = from_epiestim)
compare <- bind_rows(compare, .id = "Source")
compare <- tidyr::gather(compare,
                         key = "variable",
                         value = "value", -c(Source, rowname))

compare$rowname %<>% factor(levels = c("R[1]", "R[2]",
                                       "R[3]", "R[4]",
                                       "R[5]", "R[6]",
                                       "R[7]", "R[8]",
                                       "R[9]", "R[10]",
                                       "R[11]", "R[12]"))
                                       

ggplot(compare, aes(rowname, value, color = Source)) +
    geom_point(position = "jitter") +
    facet_wrap(~variable) +
    theme_minimal()


```



## Using incidence from Liberia

Use the entire dataset.

```{r}
lbr_incid   %<>% `[`(1:329) ## only non-NAs 
time_window <- 49
windows_end <- seq(time_window, length(lbr_incid), by = time_window)
num_windows <- length(windows_end)
T <- windows_end[num_windows]
I <- lbr_incid[seq(T)]

SI_len <- length(SI_Distr)
SI <- if(T < SI_len) SI_Distr[1:T] else c(SI_Distr, rep(0, T - SI_len))
  
sim_data <- list(T = T,
                 I = I,
                 SI = SI,
                 num_windows = num_windows,
                 windows_end = windows_end)




```

And get the ball rolling

```{r}

fit1 <- stan(
  file = "R/gravitymodel.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 5,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500          # show progress every 'refresh' iterations
)

print(fit1, pars=c("R"), probs=c(.1,.5,.9))
traceplot(fit1, pars = c("R"), inc_warmup = TRUE, nrow = 5)

lbr_stan <- summary(fit1)
lbr_stan <- lbr_stan$summary
lbr_stan <- lbr_stan[-7, c("mean", "sd", "2.5%", "25%",  "50%", "75%", "97.5%")]


```

## Estimate R using EpiEstim

```{r restim, eval = TRUE}
windows_start <- seq(1, by = time_window, length.out = length(windows_end))
r.estim   <- EpiEstim::EstimateR(I,
                                 T.Start = windows_start ,
                                 T.End = windows_end,
                                 method = "NonParametricSI",
                                 SI.Distr = SI_Distr,
                                 plot = FALSE ,
                                 CV.Posterior = 1 ,
                                 Mean.Prior = 1 ,
                                 Std.Prior = 0.5)

lbr_epiestim <- r.estim$R[, c("Mean(R)", "Std(R)",
                               "Quantile.0.025(R)",
                               "Quantile.0.25(R)",
                               "Median(R)",
                               "Quantile.0.75(R)",
                               "Quantile.0.975(R)")]

rownames(lbr_epiestim) <- rownames(lbr_stan)
colnames(lbr_epiestim) <- colnames(lbr_stan)


lbr_stan     <- tibble::rownames_to_column(data.frame(lbr_stan))
lbr_epiestim <- tibble::rownames_to_column(data.frame(lbr_epiestim))

```
And now compare them visually.

```{r}
compare <- list(stan = lbr_stan, epiestim = lbr_epiestim)
compare <- bind_rows(compare, .id = "Source")
compare <- tidyr::gather(compare,
                         key = "variable",
                         value = "value", -c(Source, rowname))

compare$rowname %<>% factor(levels = c("R[1]", "R[2]",
                                       "R[3]", "R[4]",
                                       "R[5]", "R[6]",
                                       "R[7]", "R[8]",
                                       "R[9]", "R[10]",
                                       "R[11]", "R[12]"))
                                       

ggplot(compare, aes(rowname, value, color = Source)) +
    geom_point(position = "jitter") +
    facet_wrap(~variable) +
    theme_minimal()


```
## Using incidence from Sierra Leone

Use the entire dataset.

```{r}

time_window <- 49
windows_end <- seq(time_window, length(sl_incid), by = time_window)
num_windows <- length(windows_end)
T <- windows_end[num_windows]
I <- sl_incid[seq(T)]

SI_len <- length(SI_Distr)
SI <- if(T < SI_len) SI_Distr[1:T] else c(SI_Distr, rep(0, T - SI_len))
  
sim_data <- list(T = T,
                 I = I,
                 SI = SI,
                 num_windows = num_windows,
                 windows_end = windows_end)




```

And get the ball rolling

```{r}

fit1 <- stan(
  file = "R/gravitymodel.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 5,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500          # show progress every 'refresh' iterations
)

print(fit1, pars=c("R"), probs=c(.1,.5,.9))
traceplot(fit1, pars = c("R"), inc_warmup = TRUE, nrow = 5)

sl_stan <- summary(fit1)
sl_stan <- sl_stan$summary
sl_stan <- sl_stan[-13, c("mean", "sd", "2.5%", "25%",  "50%", "75%", "97.5%")]


```

## Estimate R using EpiEstim

```{r restim, eval = TRUE}
windows_start <- seq(1, by = time_window, length.out = length(windows_end))
r.estim   <- EpiEstim::EstimateR(I,
                                 T.Start = windows_start ,
                                 T.End = windows_end,
                                 method = "NonParametricSI",
                                 SI.Distr = SI_Distr,
                                 plot = FALSE ,
                                 CV.Posterior = 1 ,
                                 Mean.Prior = 1 ,
                                 Std.Prior = 0.5)

sl_epiestim <- r.estim$R[, c("Mean(R)", "Std(R)",
                               "Quantile.0.025(R)",
                               "Quantile.0.25(R)",
                               "Median(R)",
                               "Quantile.0.75(R)",
                               "Quantile.0.975(R)")]

rownames(sl_epiestim) <- rownames(sl_stan)
colnames(sl_epiestim) <- colnames(sl_stan)


sl_stan     <- tibble::rownames_to_column(data.frame(sl_stan))
sl_epiestim <- tibble::rownames_to_column(data.frame(sl_epiestim))

```
And now compare them visually.

```{r}
compare <- list(stan = sl_stan, epiestim = sl_epiestim)
compare <- bind_rows(compare, .id = "Source")
compare <- tidyr::gather(compare,
                         key = "variable",
                         value = "value", -c(Source, rowname))

compare$rowname %<>% factor(levels = c("R[1]", "R[2]",
                                       "R[3]", "R[4]",
                                       "R[5]", "R[6]",
                                       "R[7]", "R[8]",
                                       "R[9]", "R[10]",
                                       "R[11]", "R[12]"))
                                       

ggplot(compare, aes(rowname, value, color = Source)) +
    geom_point(position = "jitter") +
    facet_wrap(~variable) +
    theme_minimal()


```

# Multiple locations

## Model

Incidence at location $i$ at time $t$ is given by
\[
  I_{i, t} \sim Poisson(\lambda_{i, t}),
\]
where 
\[
 \lambda_{i, t} = \sum_{i = 1}^{n}{R_{i, t}
                                   \sum_{s = 1}^{t}{I_{i, s} 
								   \omega_{t - s + 1}}}.
\]

## Simulated Data

### Example 1

```{r}
n.days <- 7
I0  <- matrix(c(100, 200, 300), nrow = 1)
R   <- matrix(c(1.1, 1.5, 1.7), nrow = nrow(I0) + n.days,
                                ncol = 3, byrow = TRUE)
si  <- SI_Distr
pij <- matrix(1, nrow = 3, ncol = 3)


I <- project2(incid = I0, R = R, si = si, pij = pij, n.days = n.days)

```
Preparing data for feeding into Stan.

```{r}

I  <- rbind( I0, I)
T  <- nrow(I)
N  <- ncol(I)

SI <- SI_Distr[ 1:( T + 1)]
rindex <- matrix(c(1, 2, 3), nrow = T, ncol = N, byrow = TRUE)
num_Rjt <- 3

sim_data <- list(T = T, N = N, I = I, SI = SI, rindex = rindex,
                 num_Rjt = num_Rjt)

fit1 <- stan(
  file = "R/multiple_location.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 5,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500          # show progress every 'refresh' iterations
)

print(fit1, pars=c("R"), probs=c(.01, .1,.5,.9))

```


### Example 2

```{r}
n.days <- 21
I0  <- matrix(c(100, 200, 300), nrow = 1)
R1  <- matrix(c(1.1, 1.5, 1.7), nrow = nrow(I0) + 14,
              ncol = 3, byrow = TRUE)
R2  <- matrix(c(1, 1.1, 1.2), nrow = 7,
              ncol = 3, byrow = TRUE)

R   <- rbind(R1, R2) 
si  <- SI_Distr
pij <- matrix(1, nrow = 3, ncol = 3)


I <- project2(incid = I0, R = R, si = si, pij = pij,
              n.days = n.days)

```
Preparing data for feeding into Stan.

```{r}

I  <- rbind( I0, I)
T  <- nrow(I)
N  <- ncol(I)

SI <- SI_Distr[ 1:( T + 1)]
#rindex1 <- c(rep(1, 18), rep(2, 18))
#rindex2 <- rindex1 + 2
                                        #rindex3 <- rindex2 + 2

rindex1 <- matrix(c(1, 2, 3), nrow = 14, ncol = 3, byrow = T)
rindex2 <- matrix(c(4, 5, 6), nrow = 8, ncol = 3, byrow = T)
rindex  <- rbind(rindex1, rindex2) 
num_Rjt <- 6

sim_data <- list(T = T, N = N, I = I, SI = SI, rindex = rindex,
                 num_Rjt = num_Rjt)

fit1 <- stan(
  file = "R/multiple_location.stan",  # Stan program
  data = sim_data,    # named list of data
  chains = 5,             # number of Markov chains
  warmup = 1000,          # number of warmup iterations per chain
  iter = 5000,            # total number of iterations per chain
  cores = 2,              # number of cores (using 2 just for the vignette)
  refresh = 500)          # show progress every 'refresh' iterations

print(fit1, pars=c("R"), probs=c(.1,.5,.9))

list_of_draws <- extract(fit1)
r_samples <- list_of_draws[["R"]]

p <- ggplot() + geom_histogram(aes(r_samples[, 1])) +
    geom_vline(xintercept = 1.1, col = 'red')
ggsave("1.pdf", p)
p <- ggplot() + geom_histogram(aes(r_samples[, 2])) +
    geom_vline(xintercept = 1.5, col = 'red')
ggsave("2.pdf", p)
p <- ggplot() + geom_histogram(aes(r_samples[, 3])) +
    geom_vline(xintercept = 1.7, col = 'red')
ggsave("3.pdf", p)
p <- ggplot() + geom_histogram(aes(r_samples[, 4])) +
    geom_vline(xintercept = 1.0, col = 'red')
ggsave("4.pdf", p)
p <- ggplot() + geom_histogram(aes(r_samples[, 5])) +
    geom_vline(xintercept = 1.1, col = 'red')
ggsave("5.pdf", p)
p <- ggplot() + geom_histogram(aes(r_samples[, 6])) +
           geom_vline(xintercept = 1.2, col = 'red')
ggsave("6.pdf", p)

```


