---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Estimation of Model Parameters
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---


```{r}
library(magrittr)
library(ggthemes)
library(ggplot2)
library(dplyr)
library(rstan)
library(EpiEstim)
devtools::load_all()

```


# Parameter estimation for simulated data

Load in the data and prepare to feed into Stan.

```{r ebola_params, eval = TRUE}

mean_SI  <- 14.2
CV_SI    <- 9.6 / 14.2
SItrunc  <- 40
SI_Distr <- sapply(0:SItrunc,
                   function(e) DiscrSI(e,
                                       mean_SI,
                                       mean_SI * CV_SI))
SI_Distr <- SI_Distr / sum(SI_Distr)


```

# Multiple locations

## Model

Incidence at location $i$ at time $t$ is given by
\[
  I_{i, t} \sim Poisson(\lambda_{i, t}),
\]
where 
\[
 \lambda_{i, t} = \sum_{i = 1}^{n}{p_{i, j}R_{i, t}
                                   \sum_{s = 1}^{t}{I_{i, s} 
								   \omega_{t - s + 1}}}.
\]

## Simulated Data

### Example 1

```{r}
n_loc  <- 6
n_days <- 120
I0     <- matrix(sample(10:100, n_loc, replace = T),
                 nrow = 1)


change_at <- c(31, 61, 91)
#change_at <- seq(from = 29, to = n_days, by = 28)
set.seed(42)
Rjt1 <- runif(n_loc, min = 2, max = 3) 
Rjt2 <- runif(n_loc, min = 1, max = 2)
Rjt3 <- runif(n_loc, min = 1, max = 2)
Rjt4 <- runif(n_loc, min = 0, max = 1)
Rjt  <- c(Rjt1, Rjt2, Rjt3, Rjt4)
R_sim  <- makeRmatrix( Rjt,
                       ncol = n_loc,
                       nrow = n_days + nrow(I0),
                       change_at = change_at)

#pij <- matrix(c(0.9, 0.06, 0.08,
#                0.03, 0.9, 0.02,
#                0.07, 0.04, 0.9), nrow = 3, ncol = 3)

pij <- runif(n_loc * n_loc)
pij <- matrix(pij, nrow = n_loc)
pij <- pij/rowSums(pij)

I <- project2(incid = I0, R = R_sim, si = SI_Distr,
              pij = pij,
              n.days = n_days)

```
Preparing data for feeding into Stan. 

```{r}

I  <- rbind( I0, I)
T  <- nrow(I)
N  <- ncol(I)

SI <- SI_Distr[ 1:( T + 1)]
SI[ which( is.na(SI))] <- 0

change_at <- seq(from = 14, to = n_days, by = 14) 
num_Rjt   <-  n_loc * (length( change_at) + 1)
rindex    <- makeRmatrix(seq_len(num_Rjt), nrow = T, ncol = N,
                      change_at = change_at)

sim_data <- list(T = T, N = N, I = I, SI = SI,
                 rindex = rindex,
                 num_Rjt = num_Rjt, pstay = pij)

model_file <- here::here("R", "multiple_location.stan")
fit1 <- stan(
  file = model_file,  
  data = sim_data,   
  chains = 2,        
  warmup = 1000,     
  iter = 2000,       
  cores = 2,         
  refresh = 500)     

print(fit1, pars=c("R"), probs=c(.01, .1,.5,.9))

```

## Test model fit

The fitted model has draws from the posterior distribution.
We can use these samples to project forward.

```{r test_fit, eval = TRUE}

list_of_draws <- rstan::extract(fit1)
r_samples <- list_of_draws[["R"]]
dates <- seq(from = as.Date("2018-02-07"), by = "1 day",
             length.out = nrow(I))

daily_projections <- plyr::alply(r_samples,
                                 1, 
                                 function(r.t){
                                     R_posterior <-
                                         makeRmatrix(r.t,
                                                     nrow = nrow(I),
                                                     ncol = n_loc,
                                                     change_at = change_at)                                     
                                     out <- project2(I0,
                                                     R_posterior,
                                                     SI_Distr,
                                                     pij,
                                                     n_days) 
                                     projected <- rbind(I0,
                                                        out)
                                     projected %<>%
                                      as.data.frame(.) 
                                     
                                     projected$Date <- dates
                                     return(projected)})



tmp <- bind_rows(daily_projections)
projections_distr <- projection_quantiles(tmp)


I_df <- as.data.frame(I) 
I_df$Date <- dates
I_df %<>% tidyr::gather("Country", "Incidence", -Date)


I_df$Date %<>% as.Date
p   <- ggplot(I2, aes(Date, Incidence)) +
           geom_point(size = 2, stroke = 0, shape = 16) +
           facet_wrap(~Country, scales = "free_y")
p <- p + geom_line(data = p2,
                   aes(x = Date, y = y, group = 1),
                     color = 'black', size = 0.9)
p <- p + geom_ribbon(data = p2, aes(x = Date,
                                                   ymin = ymin,
                                                   ymax = ymax,
                                                   group = 1),
                     inherit.aes = FALSE,
                     alpha = 0.5)

outname <- paste0(Sys.Date(), "-", N, "-", T, ".pdf")
outfile <- here::here("output/figures", outname)
ggsave(outfile, p)


```


```{r test_sim, eval = FALSE}
I_df <- as.data.frame(I)
I_df %<>% tibble::rownames_to_column(.)
I_tall <- tidyr::gather(I_df, var, val, -rowname)
I_tall$rowname %<>% as.numeric
ggplot(I_tall, aes(rowname, val)) + geom_point() +
    facet_grid(var~.) +
    geom_vline(xintercept = change_at)

```
