---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Estimation of Model Parameters
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  pow_dist : 2
  t.proj : 300
  n.sim : 5000
  n.dates.sim : 49
  time_window: 49
  K: 1
  pow_N_from: 1
  pow_N_to: 1
---


```{r setup, eval = TRUE}
library(magrittr)
library(ggthemes)
library(ggplot2)
library(dplyr)
library(rstan)
library(EpiEstim)
devtools::load_all()

simulated_data <- FALSE

```

# Multiple locations with full spatial model

The parameters we are now estimating are $p_{stay}$, $R_t^i$ and $\gamma$.



## User supplied parameters

```{r}
n_days      <- params$n.dates.sim
n_samples   <- params$n.sim
time_window <- params$time_window
t.proj      <- params$t.proj

```
## Gravity model parameters

Parameters that are not being estimated.

```{r gm_params, eval = TRUE}

K          <- params$K
pow_N_from <- params$pow_N_from
pow_N_to   <- params$pow_N_to

```
If data are being simulated, we need these as well.

```{r, eval = simulated_data}

pow_dist <- params$pow_dist
p.stay   <- params$p.stay

```


## Ebola parameters



```{r ebola_params, eval = TRUE}

mean_SI  <- 14.2
CV_SI    <- 9.6 / 14.2
SItrunc  <- 40
SI_Distr <- sapply(0:SItrunc,
                   function(e) DiscrSI(e,
                                       mean_SI,
                                       mean_SI * CV_SI))
SI_Distr <- SI_Distr / sum(SI_Distr)


```


## Centroids and populations 

We will use information for actual places even when using simulated
data.


```{r}

wafrica <- c("Liberia", "Sierra Leone", "Guinea")
adm0_centroids <- here::here("data", "Geography/GravityModel/raw/adm0_centroids.tsv") %>%
                   read.csv(stringsAsFactors = FALSE, sep = "\t", header = FALSE) %>%
                   filter(V1 %in% wafrica)
names(adm0_centroids) <- c("country", "id", "lon", "lat", "pop")

distances <- geosphere::distm(cbind( adm0_centroids[, "lon"],
                                     adm0_centroids[, "lat"]))

```


## Simulate Data


```{r sim_data, eval = simulated_data}
n_loc  <- 3
n_days <- 60
I0     <- matrix(sample(10:100, n_loc, replace = T),
                 nrow = 1)


change_at <- 31
#change_at <- seq(from = 29, to = n_days, by = 28)
set.seed(42)
#Rjt1 <- runif(n_loc, min = 2, max = 3) 
Rjt2 <- runif(n_loc, min = 1, max = 2)
#Rjt3 <- runif(n_loc, min = 1, max = 2)
Rjt4 <- runif(n_loc, min = 0, max = 1)
Rjt  <- c(Rjt2, Rjt4)
R_sim  <- makeRmatrix( Rjt,
                       ncol = n_loc,
                       nrow = n_days + nrow(I0),
                       change_at = change_at)

#pij <- matrix(c(0.9, 0.06, 0.08,
#                0.03, 0.9, 0.02,
#                0.07, 0.04, 0.9), nrow = 3, ncol = 3)
```

Simulate movement probability matrix.

```{r pij_sim, eval = simulated_data}

flow.matrix <- flow_matrix(longitude = adm0_centroids[, "lon"],
                           latitude  = adm0_centroids[, "lat"],
                           population = adm0_centroids[, "pop"],
                           place.names = adm0_centroids[, "country"],
                           model = "gravity",
                           K = K, pow_N_from = pow_N_from,
                           pow_N_to = pow_N_to, pow_dist = pow_dist)


## Relative risk
relative.risk <- flow.matrix / rowSums(flow.matrix, na.rm=TRUE)



pij <- probability_movement(relative.risk, p.stay)


```


```{r sim_data2, eval = simulated_data}
## matrix characterising the population movement between geographical units

I <- project2(incid = I0, R = R_sim, si = SI_Distr,
              pij = pij,
              n.days = n_days)

```

```{r sim_data_collect, eval = simulated_data}

I     <- rbind( I0, I)
dates <- seq(from = Sys.Date(),
             length.out = nrow(I),
             by = "1 day")
T  <- nrow(I)
N  <- ncol(I)

SI <- SI_Distr[ 1:( T + 1)]
SI[ which( is.na(SI))] <- 0

change_at <- seq(from = 14, to = n_days, by = time_window)

num_Rjt   <-  n_loc * (length( change_at) + 1)
rindex    <- makeRmatrix(seq_len(num_Rjt), nrow = T, ncol = N,
                         change_at = change_at)


```

## Real data

```{r who_data, eval = TRUE}

## who_wide <- here::here("data/CaseCounts/processed",
##                       "WHO_bycountry_wide.csv") %>%
##            readr::read_csv(.) %>%
##            select(Date, wafrica)

who_wide <- here::here("data/CaseCounts/processed",
                      "HealthMap_Ebola_wide.csv") %>%
            readr::read_csv(.) %>%
            filter(Date > "2014-05-15") %>%
           mutate_if(is.numeric, as.integer)

extra <- nrow(who_wide) %% 7
if( extra != 0){
        warning("Number of rows is not a multiple of 7.")
        warning(paste("Ignoring last", extra, "days."))
        who_wide %<>% head(-extra)
}


I_who <- select(who_wide, -Date) %>% `[`(seq_len(t.proj + n_days),)
dates <- pull(who_wide, Date)  %>% `[`(seq_len(t.proj + n_days))
I <- I_who[seq_len(t.proj), ]
T <- t.proj
N <- ncol(I)

SI <- SI_Distr[ 1:( T + 1)]
SI[ which( is.na(SI))] <- 0

change_at <- seq(from = 1, to = T, by = time_window)
num_Rjt   <-  N * (length( change_at) + 1)
rindex    <- makeRmatrix(seq_len(num_Rjt), nrow = T, ncol = N,
                         change_at = change_at)

```
## Prepare data for Stan

Collect data for feeding into Stan into a named list.

```{r data_prep1, eval = TRUE}

stan_data <- list(T = T, N = N, I = I, SI = SI,
                 rindex     = rindex,
                 num_Rjt    = num_Rjt,
                 population =  adm0_centroids[, "pop"],
                 dist_mat = distances,
                 alpha = 1,
                 beta = 1,
                 K = 1,
                 prior_mean = 1,
                 prior_std = 0.5)                  

```

## Run Stan Model


```{r stan_run, eval = TRUE}

model_file <- here::here("R", "full_spatial_model.stan")
fit1 <- stan(
  file = model_file,  
  data = stan_data,   
  chains = 5,        
  warmup = 1000,     
  iter = 2000,       
  cores = 2,         
  refresh = 500)     

```

```{r stan_res, eval = TRUE}

print(fit1, pars=c("R"), probs=c(.01, .1,.5,.9))
print(fit1, pars=c("gamma"), probs=c(.01, .1,.5,.9))
print(fit1, pars=c("pstay"), probs=c(.01, .1,.5,.9))

```

## Test model fit

The fitted model has draws from the posterior distribution.
We can use these samples to project forward thus doing either
goodness-of-fit test or predictive modeling.

```{r test_fit, eval = TRUE}

list_of_draws <- rstan::extract(fit1)
r_samples     <- list_of_draws[["R"]]
gamma_samples <- list_of_draws[["gamma"]]
pstay_samples <- list_of_draws[["pstay"]]


index <- sample(seq_len(nrow(r_samples)), n_samples)
#index <- seq_len(n_samples)


#dates <- seq(from = as.Date("2018-02-07"), by = "1 day",
#             length.out = nrow(I))
I0    <- I[seq_len(t.proj), ]  %>% as.matrix
n_loc <- N

daily_projections <- lapply(index, function(row){
                                     r.t   <- r_samples[row, ]
                                     gamma <- gamma_samples[row]
                                     pstay <- pstay_samples[row]
                                     R_posterior <-
                                         makeRmatrix(r.t,
                                                     nrow = t.proj + n_days,
                                                     ncol = n_loc,
                                                     change_at = change_at)
                                     flowmat <- flow_matrix(longitude = adm0_centroids[, "lon"],
                                                            latitude  = adm0_centroids[, "lat"],
                                                            population = adm0_centroids[, "pop"],
                                                            place.names = adm0_centroids[, "country"],
                                                            model = "gravity",
                                                            K = K, pow_N_from = pow_N_from,
                                                            pow_N_to = pow_N_to, pow_dist = gamma)


                                     ## Relative risk
                                     rel_risk <- flowmat / rowSums(flowmat, na.rm=TRUE)

                                     ## matrix characterising the population movement between geographical units

                                     pij <- probability_movement(rel_risk, pstay)
                                     
                                     out <- project2(I0,
                                                     R_posterior,
                                                     SI_Distr,
                                                     pij,
                                                     n_days) 
                                     projected <- rbind(I0,
                                                        out)
                                     projected %<>%
                                      as.data.frame(.) 
                                     
                                     projected$Date <- dates
                                     return(projected)})


```

```{r plot_fit, eval = !simulated_data}
tmp <- bind_rows(daily_projections)
projections_distr <- projection_quantiles(tmp)


I_df      <- as.data.frame(I_who) 
I_df$Date <- dates
I_df %<>% tidyr::gather("Country", "Incidence", -Date)


I_df$Date %<>% as.Date
p   <- ggplot(I_df, aes(Date, Incidence)) +
           geom_point(size = 2, stroke = 0, shape = 16) +
           facet_grid(Country ~ ., scales = "free_y")
p <- p + geom_line(data = projections_distr,
                   aes(x = Date, y = y, group = 1),
                     color = 'black', size = 0.9)
p <- p + geom_ribbon(data = projections_distr, aes(x = Date,
                                                   ymin = ymin,
                                                   ymax = ymax,
                                                   group = 1),
                     inherit.aes = FALSE,
                     alpha = 0.5)
p <- p + geom_vline(xintercept = dates[change_at])
outname <- paste0(Sys.Date(), "-", N, "-", T, ".pdf")
outfile <- here::here("output/figures", outname)
ggsave(outfile, p)


```


```{r test_sim, eval = simulated_data}
I_df <- as.data.frame(I)
I_df %<>% tibble::rownames_to_column(.)
I_tall <- tidyr::gather(I_df, var, val, -rowname)
I_tall$rowname %<>% as.numeric
ggplot(I_tall, aes(rowname, val)) + geom_point() +
    facet_grid(var~.) +
    geom_vline(xintercept = change_at)

```


