---
output: 
    pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: 
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "`r Sys.Date()`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---


```{r setup, eval = TRUE}
library(dplyr)
library(magrittr)
library(stringr)
library(ggplot2)
devtools::load_all()
```
## Data pre-processing

1. Read in data.
```{r s1, eval = TRUE}
promed_drc <- here::here("data/CaseCounts/raw",
                         "ProMED_ebola_drc_2018_fixed_colnames.csv") %>%
    read.csv(stringsAsFactors = FALSE)

```

2. Extract the cumulative case count as a sum of suspected and
confirmed cases.

```{r s2, eval = TRUE}
cols.to.keep <- c("Location", "Country", "Disease", "Species",
                  "HealthMap.Alert.ID", "Headline", "URL",
                  "Alert.Tag", "Feed.Name", "Lon", "Lat")

## These are columns we generate ourselves later on
cols.to.keep <- c(cols.to.keep, "Date", "Cases")
case_type <- "SCC"
step_1 <- update_cases_column(promed_drc, case_type)

```

Merge duplicate alerts.

```{r dups_check, eval = TRUE}
step_2 <- merge_duplicates(step_1, cols.to.keep)
```

3. Remove outliers and interpolate missing data.

```{r intp, eval = TRUE}
step_2 <- arrange(step_2, Date)
not.na          <- which(!is.na(step_2$Cases))
cum.incidence   <- step_2[not.na, c('Date', 'Cases')]
first.row       <- step_2[1, c('Date', 'Cases')]
first.row$Date  <- first.row$Date - 1
first.row$Cases <- 0
cum.incidence <- rbind(first.row, cum.incidence)
cum.incidence <- make_monotonically_increasing(cum.incidence)
interpolated <-  interpolate_missing_data(cum.incidence)

```
4. Check the effect of each pre-processing step visually.
```{r}
list(SCC = select(step_1, Date, Cases),
     merged_duplicated = select(step_2, Date, Cases),
     made_increasing = cum.incidence,
     interpolated = interpolated) %>%
    bind_rows(.id = "step") %>%
    ggplot(aes(Date, Cases, col = step)) + geom_point()


```

5. Determine the incidence count from the cumulative case count.


```{r}
interpolated$incid <- c(0, diff(interpolated$Cases))
```

## Estimate R

Values taken from Lancet draft.

```{r ebola_params}

mean_SI     <- 15.3
CV_SI       <- 9.1 / 15.3
SItrunc     <- 40
SI_Distr    <- sapply(0:SItrunc,
                      function(e) EpiEstim::DiscrSI(e, mean_SI, mean_SI * CV_SI))
SI_Distr    <- SI_Distr / sum(SI_Distr)


```


```{r restim, eval = TRUE}
time_window <- 7
start     <- 2:(length(interpolated$Date) - time_window)
end       <- start + time_window
end.dates <- interpolated[end, "Date"]
I <- interpolated$incid
r.estim   <- EpiEstim::EstimateR(I,
                                 T.Start = start,
                                 T.End = end,
                                 method = "NonParametricSI",
                                 SI.Distr = SI_Distr,
                                 plot = FALSE,
                                 CV.Posterior = 1,
                                 Mean.Prior = 1,
                                 Std.Prior = 0.5)



```

## Samples from the posterior distribution for R



```{r rjt, eval = TRUE}
n.dates.sim <- 7
nsim   <- 1000
cutoff <- 10
shape  <- r.estim$R[cutoff, "Mean(R)"]^2 / r.estim$R[cutoff, "Std(R)"]^2
scale  <- r.estim$R[cutoff, "Std(R)"]^2 / r.estim$R[cutoff, "Mean(R)"]
rsamples <- rgamma(nsim, shape = shape, scale = scale)
rsamples <- matrix(rsamples, ncol = 1, nrow = nsim)



```

## Project

```{r}
pmovement <- matrix(1, nrow = 1)
incid <- matrix(interpolated$incid, ncol = 1)
dates_pred <-  seq(from = max(interpolated$Date) + 1,
                   length.out = n.dates.sim,
                   by = 1)

daily.projections <- plyr::alply(rsamples, 1, function(r.t){
    out   <- project(incid,
                     as.matrix(r.t),
                     SI_Distr,
                     pmovement,
                     n.dates.sim)
    out <- data.frame(out)
    out$Date <- dates_pred
    out
})

daily.projections <- bind_rows(daily.projections)
projection_quantiles(daily.projections)
```
