---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: Forward projections using WHO incidence data
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
params:
  pow_dist : 2
  t.proj : 300
  n.sim : 2
  n.dates.sim : 49
  p.stay: 0.9
  ADM0: SIERRALEONE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, fig.path = "figures/")
```


```{r setup, echo = FALSE}
library(magrittr)
library(ggplot2)
library(dplyr)
library(EpiEstim)
devtools::load_all()

pow_dist    <- params$pow_dist
t.proj      <- params$t.proj
n.sim       <- params$n.sim
n.dates.sim <- params$n.dates.sim
p.stay      <- params$p.stay
ADM0        <- params$ADM0

```


```{r data_read, eval = TRUE}
infile     <- paste0(ADM0, "_wide.csv") 
infile     <- here::here("data/CaseCounts/processed/", infile)
incid_wide <- readr::read_csv(infile)

```

Split the data into training and validation sets.

```{r data_split, eval = TRUE}

training   <- incid_wide[1:t.proj, ]
validation <- incid_wide[(t.proj + 1):(t.proj + n.dates.sim), ]

```

# Parameters for Ebola

Culled from literature.

```{r ebola_params}

mean_SI     <- 14.2
CV_SI       <- 9.6 / 14.2
SItrunc     <- 40
SI_Distr    <- sapply(0:SItrunc, function(e) EpiEstim::DiscrSI(e, mean_SI, mean_SI * CV_SI))
SI_Distr    <- SI_Distr / sum(SI_Distr)
time_window <- 7 * 7

```

# Gravity model parameters

```{r gm_params}

pow_N_to <- pow_N_from <- 1
K        <- 1

```

# Estimating the reproduction number

For the predictions, we need to estimate R only in the training window.

```{r restim, eval = TRUE}

start     <- 2:(t.proj - time_window)
end       <- start + time_window
end.dates <- training[end, "Date"]
r.estim   <- select(training, -Date)  %>%
             plyr::alply(2, .dims = TRUE,
                            function(incid) {
                              I   <- pull(incid, 1) 
                              res <- EpiEstim::EstimateR(I,
                                                         T.Start = start ,
                                                         T.End = end,
                                                         method = "NonParametricSI",
                                                         SI.Distr = SI_Distr,
                                                         plot = FALSE ,
                                                         CV.Posterior = 1 ,
                                                         Mean.Prior = 1 ,
                                                         Std.Prior = 0.5)
                              res$R %<>% cbind(Date = end.dates)
                              return(res$R)})

```

We use the estimated R at t.proj to be the R for the window over which
predictions are being carried out.



```{r rjt}

r.j.t  <- r.estim %>%
           lapply(function(R){
                    cutoff <- which(R$Date %in% training[t.proj, "Date"])
                    shape  <- R[cutoff, "Mean(R)"]^2 / R[cutoff, "Std(R)"]^2
                    scale  <- R[cutoff, "Std(R)"]^2 / R[cutoff, "Mean(R)"]
                    return(rgamma(n.sim, shape = shape,
                                  scale = scale))}) %>%
            data.frame(check.names = F)

```

Calculate the probability of movement between locations. There is some
extra arranging of variables here to ensure that the columns in the
incidence data and movement matrix are in the same order.

```{r pmovement, eval = TRUE}

infile <- here::here("data/Geography/GravityModel/processed/", "all_centroids.csv")
centroids <- readr::read_csv(infile) %>% 
                   filter(ADM0 == ADM0) 
districts <- data.frame(CL_DistrictRes = colnames(incid_wide)[-1])
centroids <- left_join(districts, centroids)

flow.matrix  <- flow_matrix(longitude   = centroids[, "Centroid_Lon"],
                            latitude    = centroids[, "Centroid_Lat"],
                            population  = centroids[, "Pop"],
                            place.names = centroids[, "CL_DistrictRes"],
                            model = "gravity", K = K,
                            pow_N_from = pow_N_from,
                            pow_N_to   = pow_N_to,
                            pow_dist   = pow_dist)


## Relative risk
relative.risk <- flow.matrix / rowSums(flow.matrix, na.rm=TRUE)
p.movement <- probability_movement(relative.risk, p.stay)

```


```{r prediction, eval = TRUE}

incid     <- as.matrix(select(training, -Date))
dates.all <- pull(training, Date) %>%
             c(seq(max(.) + 1, length.out = n.dates.sim, by = 1))
t.max     <- t.proj + n.dates.sim - 1


daily_projections <- plyr::alply(r.j.t, 1, function(r.t){
                                    r.t   <- as.matrix(r.t)
                                    out   <- project(incid, r.t, SI_Distr,
                                                     p.movement, n.dates.sim) 
                                    incidence.proj <- rbind(incid, out)
                                    incidence.proj %<>% data.frame(check.names = F)
                                    incidence.proj %<>% cbind(Date = dates.all)
                                    return(incidence.proj[(nrow(training) + 1):t.max, ])})


weekly_projections <- lapply(daily_projections, daily.to.weekly) %>% bind_rows(.)

```

```{r predictions_write, eval = TRUE}

projections_distr <- projection_quantiles(weekly_projections)

outfile <- paste0(ADM0, "_projections_weekly_", p.stay, "_", pow_dist, "_", t.proj, ".csv")
outfile <- here::here("output", outfile)
readr::write_csv(projections_distr, path = outfile)


projections_distr <- bind_rows(daily_projections) %>%
                     projection_quantiles

outfile <- paste0(ADM0, "_projections_daily_", p.stay, "_", pow_dist, "_", t.proj, ".csv")
outfile <- here::here("output", outfile)
readr::write_csv(projections_distr, path = outfile)


```
