---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: WHO Data Pre-processing
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=6, echo=FALSE, warning=FALSE, message=FALSE, fig.path = "figures/")
```

```{r setup, echo = FALSE}
library(magrittr)
library(ggplot2)
library(ggthemes)
library(dplyr)
devtools::load_all()

```

# Introduction

The pre-processing steps consist of turning the line count into a
daily incidence series and turning the given tall data into wide data.


The current data set has three possible values for the epidemiological
case definition - confirmed, probable and suspected.
Epidemiological case definition of probable and suspected cases
differed across countries while confirmed cases were the ones
confirmed through lab report. For the purpose of the current analysis,
one approach could be to lump all of them together. The inferred date of
onset (rather than the date reported) is used for estimation.
The columns we use are : Country, EpiCaseDef (probably),
DateOnsetInferred and CL_DistrictRes. 


```{r rawdata, eval = TRUE}


WHO_raw <- here::here("data", "CaseCounts/who/rstb20160308supp1.csv") %>%
           read.csv(colClasses = c(Country = "factor",
                                   EpiCaseDef = "character",
                                   DateOnsetInferred = "Date",
                                   CL_DistrictRes = "factor")) %>%
           select(Country, EpiCaseDef,
                  DateOnsetInferred,CL_DistrictRes) %>%
           na.omit

WHO_raw$CL_DistrictRes %<>% gsub(' ', '', .)  %<>% factor



```


## Grouping by districts and interpolating missing data

For each district in a country, add the number of records for each
date to get incidence count.

```{r bydistricts, eval = TRUE}

WHO_bydistricts <- WHO_raw %>% 
                   group_by(Country, CL_DistrictRes,
                                   DateOnsetInferred) %>%
                   summarise(incid = n())

```

Within each district, if there is a date on which no cases are
recorded, we assume that there were no cases on that date and add this
to the record. At the end of this step, the incidence time series for
each district should be a daily time series.

```{r fillup, eval = TRUE}

WHO_bydistricts %<>%
    split(.$CL_DistrictRes) %<>%
    lapply(add_0incid) %<>%
    bind_rows


WHO_bydistricts %<>% rename(Date = DateOnsetInferred)

WHO_bycountry <- WHO_bydistricts %>%
                  group_by(Country,Date) %>%
                  summarise(incid = sum(incid)) 


```

```{r who_raw_viz, eval = TRUE}

ggplot(WHO_bydistricts, aes(Date, incid)) +
    geom_point() +
    facet_wrap(~CL_DistrictRes) +
    ggtitle("Incidence in each district") +
    xlab("") + theme(axis.text.x = element_text(angle = 45, vjust = 0.2)) +
    theme_tufte() + geom_rangeframe() 

ggplot(WHO_bycountry, aes(Date, incid)) +
    geom_point() +
    facet_wrap(~Country) +
    theme_minimal() +
    ggtitle("Incidence in each country") +
    xlab("") + theme(axis.text.x = element_text(angle = 45, vjust = 0.2)) +
    theme_tufte() + geom_rangeframe()     

```

Some spelling variations between the WHO data and the data set
containing the district co-ordinates need to be taken
into account.

```{r spelling_fix, eval = TRUE}
WHO_districts$CL_DistrictRes %<>% factor
WHO_districts$CL_DistrictRes %<>%
    plyr::mapvalues(from = c("KISSIDOUGO",  "YOMOU",  "N'ZEREKORE", "GBARPOLU" ),
                    to   = c("KISSIDOUGOU", "YAMOU",  "NZEREKORE", "GBAPOLU"))

```

Write the files.

```{r who_ddump, eval = TRUE}

outfile <- here::here("data", "CaseCounts/processed/WHO_bydistricts.csv")
readr::write_csv(WHO_bydistricts, outfile)

outfile <- here::here("data", "CaseCounts/processed/WHO_bycountry.csv")
readr::write_csv(WHO_bycountry, outfile)

```

Now we will convert the data from the tall to the wide format.

```{r who_country_viz, eval = TRUE}

WHO_bydistricts_split <- WHO_bydistricts %>%
                         split(.$Country) 
lapply(WHO_bydistricts_split, function(df){
                                  country <- df$Country[1]
                                  outfile <- paste0("figures/",
                                                    country, "-incid.png")
                                  p <- ggplot(df, aes(Date, incid)) +
                                      geom_point(size = 1.1,
                                                 stroke = 0,
                                                 shape = 16) +
                                      facet_wrap(~CL_DistrictRes) +
                                      ggtitle(paste("Incidence in the districts of", country))
                                  p <- p + geom_rangeframe() + theme_tufte()
                                  ggsave(outfile, p)})    

```


```{r tall_to_wide, eval = TRUE}

lapply(WHO_bydistricts_split, function(df){
                                  df %<>% na.omit
                                  country <- df$Country[1] 
                                  country %<>% gsub(' ', '', .) %<>% toupper
                                  outfile <- here::here("data",
                                                        "CaseCounts/processed/")
                                  outfile <- paste0(outfile, country, "_wide.csv")
                                  df_wide <-  df %>% ungroup %>%
                                                     select(-Country) %>%
                                                     tidyr::spread(CL_DistrictRes,
                                                                   incid, fill = 0)
                                  readr::write_csv(df_wide, outfile)})


```

Also write out the data for all districts for later analysis.

```{r who_all, eval = TRUE}

WHO_wide <- WHO_bydistricts %>%
                    ungroup %>%
                    select(-Country) %>%
                    tidyr::spread(CL_DistrictRes, incid, fill = 0)

outfile <- here::here("data",
                      "CaseCounts/processed/WHO_wide.csv")
readr::write_csv(WHO_wide, outfile)



```
