---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
title: 
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---

```{r setup}
library(magrittr)
library(ggplot2)
library(dplyr)

```

In the WHO dataset, location of residence was recorded to the second
administrative level, termed prefecture in Guinea, county in Liberia
and district in Sierra Leone. There are minor differences between the
names recorded in the WHO dataset and as available in the dataset
recording their latitude, longitude and population. We will clean up
these differences and save the clean files.

```{r who_districts, eval = TRUE}

infile <- here::here("data/CaseCounts/processed/", "WHO_bydistricts.csv")
WHO_districts <- readr::read_csv(infile) 

```

```{r centroids_merge, eval = TRUE}
guinea_sierra <- here::here("data","Geography/GravityModel/raw/adm2-fixed.txt") %>%
                     read.csv(stringsAsFactors = FALSE,
                              sep = "\t",
                              header = TRUE) %>%
                    filter(ADM0 %in% c("Guinea", "Sierra Leone"))  

liberia <- here::here("data", "Geography/GravityModel/raw/adm1_centroids_fixed.tsv") %>%
                   read.csv(stringsAsFactors = FALSE,
                            sep = "\t",
                            header = TRUE) %>%
                   filter(ADM0 == "Liberia")  

guinea_sierra %<>%
    select(ADM0, ADM2, Centroid_Lon, Centroid_Lat, Pop) %<>%
    rename(CL_DistrictRes = ADM2)

liberia %<>%
    select(ADM0, ADM1, Centroid_Lon, Centroid_Lat, Pop) %<>%
    rename(CL_DistrictRes = ADM1)


all_centroids <- bind_rows(guinea_sierra, liberia) 


```
We now convert all non-ASCII letters to ASCII characters (e.g.,
accented e is converted to e), remove spaces and convert case.

```{r names_cleanup, eval = TRUE}
all_centroids$CL_DistrictRes     %<>%
    iconv(to='ASCII//TRANSLIT')  %<>%
    gsub("\\'", '', .)  %<>%
    gsub(' ', '', .) %<>%
    toupper %<>%
    factor

all_centroids$ADM0   %<>%
    gsub(' ', '', .) %<>%
    toupper 

```
In the WHO data-set, no distinction has been made between Western
Urban and Western Rural districts of Sierra Leone. So we treat the
co-ordinates of Western Urban as those of Western district.

```{r west_urban, eval = TRUE}
all_centroids$CL_DistrictRes %<>% gsub("WESTERNURBAN", "WESTERN", .)

```


Some of the prefectures of Guinea are not represented in the WHO
data. These are: Gaoual, Koubia, Koundara, Labe, Lelouma, Mamou
and Mandiana. Hence we will drop these from the data frame of
centroids.

```{r guinea_missing, eval = TRUE}

missing <- setdiff(unique(all_centroids$CL_DistrictRes),
                   unique(WHO_districts$CL_DistrictRes))

notmissing    <- !(all_centroids$CL_DistrictRes %in% missing)
all_centroids <- all_centroids[notmissing, ]


all_centroids$CL_DistrictRes %<>% factor


```

Both files are now consistent and we can write them out.

```{r centroids_write, eval = TRUE}

outfile <- here::here("data/Geography/GravityModel/processed/", "all_centroids.csv")
readr::write_csv(all_centroids, path = outfile)

```
