---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
title: Exploratory analysis of weekly district data from WHO 
author:
- name: Sangeeta Bhatia
  affiliation: Imperial College London
abstract: 
keywords: 
date: "Y"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
spacing: double
bibliography: 
biblio-style: apsr
endnote: no
---

```{r setup, eval = TRUE}
library(dplyr)
library(magrittr)
library(stringr)
library(ggplot2)
library(ggthemes)
```
# Data load and comparison with data published in Tini's paper

The two different data sources are Patient Database and Situation
Report. According to the website, Patient Database ``provides
information on the symptoms, diagnosis and outcomes of treatment for 
each suspected, probable or confirmed Ebola patient; typically gives 
the fullest representation of the epidemic, but is often incomplete 
for the most recent days or weeks.'' 

Situation Report ``Contains daily
district information on number of suspected, probable or confirmed
Ebola patients; lacks many of the details found in the patient 
database, but provides the best data on the status of the epidemic 
now and for recent days.''

```{r}
sl_website <- here::here(
  "data/CaseCounts/processed",
  "processed_who_sl_2015-08-19.csv"
) %>%
  read.csv(.)
sl_website$week_of_year <- factor(sl_website$week_of_year)
```

And WHO data from Tini's paper.
```{r}
sl_linelist <- here::here(
  "data/CaseCounts/who",
  "rstb20160308supp1.csv"
) %>%
  read.csv(colClasses = c(
    Country = "factor",
    EpiCaseDef = "character",
    DateOnsetInferred = "Date"
  )) %>%
  select(
    Country,
    EpiCaseDef,
    DateOnsetInferred,
    CL_DistrictRes
  ) %>%
  rename(location = CL_DistrictRes) %>%
filter(Country == "Sierra Leone")

sl_linelist <- filter(sl_linelist, !is.na(DateOnsetInferred))
sl_linelist <- arrange(sl_linelist, DateOnsetInferred)
sl_weeks <- lubridate::week(sl_linelist$DateOnsetInferred) %>%
  stringr::str_pad(2, pad = "0")

sl_linelist$week_of_year <- paste0(
  lubridate::year(sl_linelist$DateOnsetInferred),
  "-W",
  sl_weeks
)
```

Are the dates/weeks missing from the above linelist the same as the dates
for which data from website contains blanks?



```{r}
all_dates <- seq(
  from = min(
    sl_linelist$DateOnsetInferred,
    na.rm = TRUE
  ),
  to = max(
    sl_linelist$DateOnsetInferred,
    na.rm = TRUE
  ),
  by = "1 day"
)

dates_na_linelist <- all_dates[!(all_dates %in%
  sl_linelist$DateOnsetInferred)]

weeks_padded <- stringr::str_pad(
  lubridate::week(dates_na_linelist),
  2,
  pad = "0"
)
weeks_na_linelist <- paste0(
  lubridate::year(dates_na_linelist),
  "-W",
  weeks_padded
)
```

And for data from the website.


```{r}
idx <- which(is.na(sl_website$new_cases))
weeks_na_website <- sl_website$week_of_year[idx] %>% unique()
```

What weeks are NA for the linelist but not in website data?

```{r}
setdiff(weeks_na_linelist, weeks_na_website)
```
Weeks from 2013. We won't worry about this.
What weeks are NA for the website data but not in linelist?

```{r}
setdiff(weeks_na_website, weeks_na_linelist)
```

Dig a little deeper. For each district, for each case type, where we
have a blank in the website, what info do we have in the linelist?

```{r}
by_district <- filter(sl_website, is.na(new_cases)) %>%
  split(list(.$location, .$case))

tmp <- lapply(by_district, function(x) {
  district <- unique(x$location)
  filter(sl_linelist, CL_DistrictRes == district &
    week_of_year %in% unique(x$week_of_year))
})
```

Let us treat NAs as 0 and compare weekly data.


```{r}
idx <- which(is.na(sl_website$new_cases))
sl_website$new_cases[idx] <- 0
```
Add confirmed and probable cases.

```{r}
sl_website_total <- group_by(
  sl_website,
  week_of_year,
  data_source,
  location
) %>%
  summarise(total_cases = sum(new_cases))
```

Linelist to weekly

```{r}
sl_linelist_weekly <- select(
  sl_linelist, -Country,
  -EpiCaseDef, -DateOnsetInferred
) %>%
  group_by(location, week_of_year) %>%
  mutate(total_cases = n())

sl_linelist_weekly$data_source <- "linelist"
sl_linelist_weekly <- select(
  sl_linelist_weekly,
  week_of_year,
  data_source,
  location,
  total_cases
)
```

```{r}

all_sources <- rbind(sl_linelist_weekly, sl_website_total)
all_sources$week_of_year <- factor(all_sources$week_of_year)
all_sources %>%
    split(.$location) %>%
    lapply(function(x){
        p <- ggplot(x, aes(week_of_year,
                           total_cases,
                           col = data_source)) + geom_point()
        p <- p + theme_bw()
        p <- p + theme(axis.text.x =
                           element_text(angle = 90,
                                        hjust = 1,
                                        size = 2))
        here::here("output",
                   paste0(unique(x$location), ".png")) %>%
            ggsave(p)})

```

So there is significant discrepancy between the three data sources.
Let's just focus on the weeks where the website says blank (or 0 now).


```{r data_read, eval = TRUE}
sl_website_0s <- filter(sl_website_total, total_cases == 0) %>%
    droplevels()
sl_linelist_0s <- filter(sl_linelist_weekly,
                         week_of_year %in%
                         unique(sl_website_0s$week_of_year)) %>%
        droplevels()

rbind(sl_linelist_0s, sl_website_0s) %>%
    split(.$location) %>%
    lapply(function(x){
        p <- ggplot(x, aes(week_of_year,
                           total_cases,
                           col = data_source)) + geom_point()
        p <- p + theme_bw()
        p <- p + theme(axis.text.x =
                           element_text(angle = 90,
                                        hjust = 1,
                                        size = 2))
        here::here("output",
                   paste0("0_", unique(x$location), ".png")) %>%
            ggsave(p)})
```

